---
title: "Notes from call with Doug Kuhns"
author: "Randy Johnson"
date: "6/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following are measures and statistics that Doug has used in the past. Several of these will benefit a lot from a larger number of cells.


### Survival analysis

Modeling the proportion of cells that cross the lower threshold as a function of time.


### Chemotactic efficiency

Compare how far each cell traveled vs net vertical/directed movement. The measure we'll use for Chemotactic Efficiency, $CE$, is the vertical distance traveled divided by the total distance traveled. This will be very close to 1 for cells that move primarily in a directed fashion and much lower for cells that travel greater distances laterally.

\begin{align*}
  CE & = \frac{\mbox{Vertical distance}}{\mbox{Total distance}}
\end{align*}


### Angle of Migration

The angle of migration, $\theta$, is defined as the absolute angle of net movement from the vertical axis (see figure below). This is calculated as $$\theta = \tan^{-1}(x/y)$$ where $x$ is the net random movement (i.e. in the horizontal direction) and $y$ is the net directed movement (i.e. in the vertical direction).

```{r angle, echo=FALSE, fig.width = 2, fig.height = 2, fig.align='center'}
set.seed(293847)
# cell track
len = 300
x <- cumsum(rnorm(len, .05))
y <- -(1:300)#cumsum(rnorm(len, -1))

# arc showing angle
x_arc <- seq(from = 0, to = 13, length = 300)
y_arc <- -sqrt(1000 - x_arc^2)

# for relative scaling and centering of x
expansion <- 5
offset <- 40

# render plot (ugly but efficient)
par(mar = rep(0,4))
plot(x + offset, y, type = 'l', xlim = range(x)*expansion, col = rgb(0,0,0,.5),
     bty = 'n', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '')
arrows(x[1] + offset, y[1], x[len] + offset, y[len], length = 0.1)

lines(x_arc/2 + offset, y_arc - 60)

text(x[1] + offset - 10, y[1] - 8, 'Start', cex = 0.5)
text(x[len] + offset + 10, y[len], 'End', cex = 0.5)
text(4 + offset, -105, expression(theta), cex = 0.75)
abline(v = x[1] + 40, lty = 2, col = 'grey50')
abline(h = -1, lty = 2, col = 'grey50')
```


### Total Velocity

Similarly, total velocity, $V$, is the net distance traveled, or the length of the vector in the image above, divided by the total number of frames the cell was tracked over: $$V = \frac{\|\langle Start, End \rangle\|_2}{\mbox{n frames}}.$$


### Additional Measures

The following are some additional measures that would be good to explore.

* Peak velocity
* Time to peak velocity
* Slope of post-peak velocity
